### 2016-11-25 Composition_Ecology.Rmd
More text updates and reviewr comments.
Information about detectability & heterogeneity differences between FIA & PLS, some simple analysis of climate normals, added a whole chunk of supplementary code to look at how choices of normals would affect the envelopes.

Added some references focused on historical temperature estimates.

Changed the structure of the `clim_boxplot()` function so it accepts filenames as an argument.  I really should change `confounding_table` so it's a function, then I could pass in the difference `vegclim_table` alternates.

---
bibliography: goringetal_references.bib
csl: ecology.csl
output:
  word_document: 
    reference_docx: word_template.docx  
  pdf_document:
    pandoc_args: -V geometry:vmargin=1in -V geometry:hmargin=1in
  md_document:
    variant: markdown_github
pandoc_args: -H margins.sty
---

<!--

```{r setup-knitr, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = " ",
  error = TRUE,
  eval = TRUE,
  cache = TRUE,
  tidy = TRUE

)

library("pander")
library("plyr")

panderOptions('missing', '--')
panderOptions('digits', 4)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.continues', NULL)
pander::panderOptions('table.style', 'rmarkdown')
pander::panderOptions('table.split.table', 300)
grDevices::ps.options(fonts = 'serif')

```
-->

# Effect of historical land-use and climate change on tree-climate relationships in the north-central United States

**Statement of Authorship**

* Simon J Goring: Co-conceived paper, performed analysis, developed figures and co-wrote the paper
* John W Williams: Co-conceived paper, and co-wrote the paper

**Data Accessibility**

* All data for this manuscript, along with code and version-control are available at http://github.com/SimonGoring/CompositionClimate.  In addition, an archived copy of the submitted version is available with the following DOI: XXXX

**Authors**

* Simon J Goring, Department of Geography, University of Wisconsin, Madison, 550 N Park St, Madison WI 53706
* John W Williams, Department of Geography, University of Wisconsin, Madison, 550 N Park St, Madison WI 53706; Center for Climatic Research, University of Wisconsin, Madison, 1225 W Dayton St, Madison WI 53706

**Keywords**

* Anthropocene, climate change, climate disequillibrium, forest inventory and analysis (FIA), fundamental niche, historical ecology, land use, niche shift, Public Land Survey System, realized niche

**Running Title**

* Causes of recent climate shifts in trees

**Article Type**: Letters

**Number of Words**

* Main Text: 4900
* Text boxes: 0

**Number of References**: 77

**Number of Figures, Tables & Text Boxes**: 6 (5 figures, 1 table)

**Corresponding Author**

* Simon J Goring
  * Department of Geography, University of Wisconsin, Madison, 550 N Park St, Madison WI 53706, goring@wisc.edu

# Abstract

(143 words)

Contemporary forest inventory data are widely used to understand environmental controls on tree species distributions and to project forest responses to climate change, but the stability and representativeness of contemporary tree-climate distributions is poorly understood. Here we show that tree-climate relationships for 15 common tree genera in the north-central US have significantly altered over the last two centuries due to historical land use, pathogen outbreaks, and climate change. Realized niches have shifted towards higher minimum temperatures, cooler maximum temperatures, and higher rainfall. A new attribution method implicates both historical climate change and land use in these shifts, with the relative importance varying among genera and climate variables. Most climate/land-use interactions are compounding, in which historical land-use reinforces climatic shifts in species distributions toward cooler or wetter distributions. The prevalence of compounding interactions imply that contemporary-based models of species distributions may underestimate species resilience to climate change.

/pagebreak

```{r, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Load climate data and generate the pre-sett and modern bounds:
#  Pre-settlement climate: 1900 - 1930
#  Modern climate: 1960 - 1990

source("R/Setup_R.R")
source("R/load_climate.R") # gives `climate` with elements `then`, `now` and `names`
source("R/load_veg.R")

best.taxa <- c("beech", "ironwood", "hemlock", "fir", "cedar.juniper",
              "basswood", "spruce", "elm", "ash", "pine", "oak",
              "tamarack", "maple", "poplar", "birch")

pls_data <- get_pls("0.9-6")
agg_dens <- get_fia()[,colnames(pls_data)]

land_use <- land_use[land_use$cell %in% c(pls_data$cell, agg_dens$cell),]

source("R/natural_earth.R")

natural <- natural_earth()

```

# Introduction

Contemporary species occurrence databases, and, increasingly, paleohistorical datasets are used to map distributions of species and species traits within environmental space, study underlying processes, and prepare for climate change [@lamanna2014functional;@moritz2013future;@maguire2015modeling;@ivory2016niche;@pearman2008niche]. These distributions of species in geographic and environmental space, combined with future climate scenarios, can be used to assess species exposure and sensitivity to climate change, identify species and habitats of concern, set conservation priorities, and prepare for climate-driven shifts in habitat suitability [@pellatt2012using;@iverson2013tree;@iverson1998predicting;@anderson2013nichemodels;@guisan2013predicting]. Forest inventory datasets, such as the US Forest Inventory Analysis (FIA), are particularly rich observational datasets that have been used to assess early signals of range shifts via the distributions of mature trees and seedlings [@monleon2015evidence;@zhu2012failure] and to parameterize models of current and future tree distributions [@iverson2008estimating;@iverson1998predicting].

Increasingly, the predictive capacity of ecological models is being improved by integrating contemporary observational data with paleohistorical data, when available [@moritz2013future;@maguire2015modeling]. Projections relying only on contemporary observational data may overlook plasticity in species-climate responses [@maiorano2013building], incompletely delimit species fundamental niches [@jackson2000responses], miss responses of species to past no-analog climates [@veloz2012no], or fail to detect effects of anthropogenic land use, and other historical legacies [@svenning2015influence]. Paleoecological records from the last deglaciation show that the time lag of forest responses to abrupt climate change varies from a few years to centuries [@williamsinpress;@webb1986vegetation;@ordonez2013realized;@svenning2015influence], suggesting that tree distributions have yet to fully respond to recent climate changes. In principle, mechanistic dynamic vegetation models [*e.g.*, @leiblein2016modelling] can represent the physiological, demographic, and dispersal processes that cause tree populations to lag climate change.  In practice, however, the parameterization of these models is poorly constrained for processes operating at large temporal and spatial scales [@moorcroft2006close;@matthes2016benchmarking].

Historical land use has substantially affected species distributions and vegetation structure [*e.g.*, @gehrig2007tree] and may cause species distributions to incompletely fill available climates, or otherwise be at disequilibrium [@early2014climatic;@sarmento2013impacts;@ivory2016niche]. Anthropogenic land use is structured and is often biased within climate space, hence, species distributions and protected natural areas often represent a biased selection from the available environmental space [@scott2001nature]. Biasing can confound studies that use contemporary distributional data to understand the environmental controls on species distributions and predict species responses to future climate scenarios [@pyke2004habitat].

Interactions between climate and land-use change can be classified into three types: compounding, confounding and counteracting [modified from @pyke2004habitat]. Compounding effects occur when climatic and land-use changes cause species distributions to shift in the same direction along an environmental gradient.  For example, warming has been compounded by land-use conversion at the lowest elevations in the northern Sierra Nevada Mountains of California, both causing upslope distributional shifts for butterfly species [@forister2010compounded]. Confounding effects may cause little overall change on species distributions, with land-use pressure occurring across the species' environmental gradient. Predictive models for Garry oak (*Quercus garryana*) using only contemporary data [@pellatt2012using] may be confounded by land-use conversion that excludes Garry oak from the deep-soil sites it formerly occupied [@macdougall2004defining;@pellatt2014environmental], but land use has not systematically shifted Garry oak distributions in climate space. Counteracting effects occur when climate and land use change act in opposite directions, narrowing the available geographic and climatic space available for species. For example, bird species in the Sierra Nevadas are expected to move downslope due to increasing precipitation [@tingley2012push], but this downslope migration is limited by land-use conversion in the lowlands.

```{r figure1, echo = FALSE, warning = FALSE, message=FALSE, fig.width=6, fig.height=5}

land_use$class[land_use$class == "Water"] <- NA

ext <- c(-100000, 1050000, 600000, 1600000)

land_use_plot <- natural[[2]] + 
  geom_tile(data = land_use, aes(x = x, y = y, fill = class)) +
  coord_equal(xlim = ext[1:2], 
              ylim = ext[3:4]) + 
  scale_fill_manual(values = c("#ff7f00", "#e41a1c", "#33a02c")) +
  theme(legend.position = "none")

source("R/figures/climate_ts.R")

climate_values <- climate_ts(unit.raster, pls_data, climate$now, climate$then)

clim_change <- climate_values$modern$value - climate_values$past$value

ggsave(plot = arrangeGrob(climate_values$plot, land_use_plot, nrow = 1, widths = c(3, 4)), 
       filename = "figures/figure_one.tiff")

```
![](figures/Fig_1_Composite.png)

`r fig_nums("land_use_clim", "*(a-e) Climate change in the Upper Midwest over the last two centuries using PRISM data and North American Drought Atlas PDSI reconstructions. Modern climate shows lower $T_{max}$, but higher $T_{min}$ relative to early-19^th^ Century normals, along with increasing $P_{ann}$.  PDSI shows strong coherence with $P_{ann}$ in this region. (f) Patterns of land-use change in the upper Midwestern United States, with the 'Tension Zone', differentiating sub-boreal from southern broadleafed forests superimposed. Land use data from the NLCD [@jin2013comprehensive].*")`

Tree distributions and community composition in the upper Midwestern US have been heavily influenced by land use conversion associated with EuroAmerican Settlement in the 19^th^ and 20^th^ Centuries [@goring2016preset;@schulte2007homogenization;@hanberry2012comparison;@rhemtulla2009historical], outbreaks of exotic pathogens [@barnes1979stem], climate changes accompanying the end of the Little Ice Age [@hotchkiss2007response] and current anthropogenic warming [@wicci2011climate]. Modern forests, as represented by FIA surveys [@woudenberg2010forest], show greater homogeneity, weaker ecotones, and significantly different species associations than forests surveyed as part of the Public Land Survey System (PLS) prior to widespread land-use conversion in the late 1800s [@goring2016preset;@schulte2007homogenization]. Wildland ecosystems have been converted to agricultural and urban land use [@rhemtulla2009historical], particularly in historical prairie and savanna ecosystems (`r fig_nums("land_use_clim", display = "cite")`f).  In northern forested regions, ranges of early successional species have expanded and abundances of late-successional species are reduced due to widespread logging in the 19^th^ and 20^th^ century, and ongoing land management [@schulte2007homogenization;@goring2016preset].  Additionally, the formerly dominant *Ulmus americana* has experienced rangewide population collapses caused by the introduction of exotic pathogens [@barnes1979stem], as did *Castanea dentata*, with other tree species currently threatened (*Tsuga canadensis*, *Fraxinus* spp.)

Here we demonstrate significant shifts in the climatic distributions of tree genera in the upper Midwest over the last 100 years as a result of historical climate change, land use, and pathogen outbreaks, and we diagnose the relative influence of historical climate and land use changes on these shifts.  We use gridded historical and contemporary climate observations (PRISM-LT) and estimates of pre-settlement and modern forest vegetation [@goring2016preset] to map the geographic and climatic distributions for 15 major tree taxa in the upper Midwestern United States.  We construct four observed and hypothetical sets of tree-climate relationships, using a 2x2 factorial of pre-settlement and modern vegetation and climates.  We develop a new method, based on Hellinger distances among tree-climate distributions, to quantify the total change and attribute the relative importance of historical vegetation and climate change. We assess whether climate and vegetation changes are compounding, counteracting or confounding based on the magnitude and direction of species-climate shifts. We use this framework to discuss the potential impacts of historical land-use and climate change on inferences about species resilience to climate change, as predicted by species-distribution models based soley on modern distributional data.

# Materials and Methods

## Historical and Modern Vegetation Data:

Estimates of settlement-era forest composition and tree distribution are based on survey data from the PLS [@white1983history;@schulte2001original;@goring2016preset], aggregated to a 64 km^2^grid across the upper Midwestern United States [@goring2016preset]. We briefly review the methods here. PLS data originally were collected on a 1.61 x 1.61 km (1 x 1 mi) grid across the Upper Midwest from 1830 to   1910 [@schulte2001original;@almendinger1996minnesota;@white1983history]. At each survey point surveyors noted the closest two to four trees and recorded their common name, distances from survey points, diameters at breast height, and azimuths to each tree. Goring et al. [-@goring2016preset] aggregated this data to an 8x8 km grid, resulting in a gridded data product with an average of 61 survey points per cell, or approximately 120 trees per cell. This process included data cleaning, taxonomic standardization, and the application of spatially varying correction factors [@goring2016preset]. Using this data, we produce presence-absence distributions for each taxon, across the upper Midwest for 15 common tree genera. A few common names, such as "Ironwood" can not be clearly resolved to a single tree genera and so we combine *Ostrya* and *Carpinus*, and *Thuja* and *Juniperus*. 'Presence' is defined as at least one recorded tree in any 8x8 km grid cell and 'range' is considered to be the extent in climate space of all cells with recorded presence for a specific taxon.

Modern forest data are from the US FIA program, which provides detailed inventories of contemporary tree distributions, for all regions with >10% forest cover [@woudenberg2010forest]. The FIA uses a nationally standardized sampling procedure that has evolved over time.  In the Phase 2/3 system, beginning in 1998, each plot consists of four, 7.32m radius, circular sub-plots which are sampled intensively [@woudenberg2010forest].

Comparisons between historical and modern vegetation must address differences in sampling design between the PLS and FIA datasets [@goring2016preset;@kronenfeld2014validating;@kronenfeld2010influence]. Here, the primary solution for maximizing comparability is via aggregation, both spatially, by aggregating records from both datasets to a common scale (8x8 km), and taxonomically by aggregating taxa to the level of genera. Nonetheless, differences persist.  The FIA contains fewer plots per 8 km cell, with more trees per plot than the PLS data and FIA observations are collected only from forested or partially forested landscapes [@woudenberg2010forest], causing many tree taxa to have smaller spatial extents in the FIA than in the PLS datasets. Scattered trees may be present in regions without permanent or temporary FIA plots, complicating the ability to assign 'absence' within the FIA. However, the practice of using FIA data as a surrogate for tree distribution data is widespread [@iverson2013tree;@iverson1998predicting], so these complications are general. In prior comparative analyses of the PLS and FIA data, Goring et al. [-@goring2016preset] tested the effect of differential sampling design on the mapped patterns of novel and disappearing forests, and found little effect. FIA sampling represents a system with the potential for high heterogeneity within cells because of intensive sampling at the plot level, and, with one to few plots within an 8 x 8km grid cell, high between cell heterogeneity [@goring2016preset]. PLS sampling is likely to have low heterogeneity within and between cells because of the uniform spatial distribution, and the high number of points within cells [@goring2016preset].  Detectability at the plot level for FIA data is likely to be higher than for the PLS, however, given the variable radius of PLS plots, detectability of trees at forest margins is likely to be higher in the PLS, and PLS sample plots are likely to sample higher landscape heterogeneity within a given cell because of their uniform distribution across the landscape.

For all species, the range within the Upper Midwest represents an incomplete sample of the species range, however, the prairie-forest boundary along the western margin represents a southern and western range limit for many species, and represents a climatic space that is warmer and drier than elsewhere in the region.

## Historical and Modern Climate Data

Historical and modern climate data are derived from monthly PRISM LT data for 1895 to 2014 CE [@prism2004]. The PRISM climate variables are available at 800m resolution and have been interpolated from station locations using elevation, aspect and other data.  Estimates for mean daily July temperature ($T_{max}$), mean daily January temperature ($T_{min}$), annual temperature range ($T_{diff}$: $T_{max}$ - $T_{min}$), and annually summed daily precipitation ($P_{ann}$) were resampled to the 8 x 8km grid used for PLS and FIA data. These climate variables were chosen because of their relative importance as demonstrated in prior research [@iverson1998predicting] and their use in past climate reconstruction [@maiorano2013building].  Recent statistical advances have opened the possibility of obtaining temperature and precipitation estimates from historical observations as early as the 1830s [@tiptoninreview], however the behaviour of uncertainties for derived variables such as potential evapo-transpiration, which ranked highly in previous work [@iverson1998predicting] are not currently known.

Two periods were selected for calculation of climate normals: 1895 to 1919 for the historical period and 1990 - 2014 for the modern period.  There is an unavoidable temporal mismatch between the historical climate data and the PLS survey period, on the order of 50 to 70 years; few meteorological records exist in this region prior to 1895 CE, and no gridded dataset exists at high spatio-temporal resolution for this time period for all climate variables.  The effects of temporal mismatch should be minor given that Northern Hemisphere temperature trends were generally small during the 19^th^ century [@mann2009global].  To check whether an 1895 to 1919 CE climate normal is a reasonable surrogate for pre-settlement climates we compare the instrumental record against dendroclimatic reconstructions of the Palmer Drought Severity Index [@heim2002review] (`r fig_nums("land_use_clim", display = "cite")`), and test July temperature reconstructions from observer data against our $T_{max}$ (Supplement 1).  Baker et al. [-@baker1985minnesota] show long term increases in mean annual temperature at Ft. Snelling and other spatially proximate records, while corrected estimates from Fort Leavenworth in Kansas [@burnette2010daily] show increasing temperatures in all seasons, with faster increases during winter (0.10^o^C decade^-1^) than during summer (0.07^o^C decade^-1^).  This implies that historical estimates may be biased by +0.35^o^C for $t_{max}$, by +0.5^o^C for $t_{min}$ and +0.15^o^C for $t_{diff}$.

## Shifts in Species Distributions in Climate Space

![](figures/figureFour_panels.tiff)

`r fig_nums("fourPanel", "(a) The 2x2 factorial design used to diagnose the effects of historical climate and vegetation change on observed shifts in tree-climate relationships. Observed and hypothetical tree-climate relationships for one of four combinations of pre-settlement vegetation (pink fill, left column) or modern vegetation (blue fill, right column) and pre-settlement climate (solid border, top row) or modern climate (dashed border, lower row). (b-e) Representative normalized probability densities for *Larix* for each of the four combinations of climate and vegetation states, plotted relative to $T_{max}$ (b), $T_{diff}$ (c), $T_{min}$ (d), and $P_{ann}$ (e).")`

Distributions of tree genera within climate space are described as univariate probability density functions (PDFs) for the pre-settlement and modern eras.  Differences among PDFs are quantified using Hellinger distances [@matthes2016benchmarking] for the four combined states of vegetation and climate ($V_{H}C_{H}$, $V_{H}C_{M}$, $V_{M}C_{H}$, $V_{M}C_{M}$, where V represents vegetation, C represents climate, and M and H represent modern and historical conditions). The total shift ($d_{tot}$) is calculated as the Hellinger distance between $V_{M}C_{M}$ and $V_{H}C_{H}$ (`r fig_nums("fourPanel", display = "cite")`a: top left to bottom right). The effect of historical climate change on shifting climate distributions of tree genera is $d_{c}$, the difference between PLS and FIA tree distributions, overlaid on late 19th-century climates ($V_{H}C_{H}$ versus $V_{H}C_{M}$; `r fig_nums("fourPanel", display = "cite")`a; left column). The difference between the PLSS and modern tree distributions ($d_{v}$) when overlaid on historical climate ($V_{H}C_{H}$ versus $V_{M}C_{H}$) is attributed largely to land use change, as described above, given the widespread post-settlement disturbance in the upper Midwest; Figure 2a; top row).

We develop an attribution index, ($\delta_{d} = d_{c} - d_{v}$), such that negative values of $\delta_{d}$ indicate stronger influence of historical land use on shifting climate distributions, while positive values indicate a stronger influence of historical climate change on shifting tree distributions.

To assess whether the effects of historical land use and climatic changes are compounding, counteracting or confounding we perform a set of t-tests, using the conservative Bonferroni correction to account for multiple tests.  For each taxon and climate variable pair we test whether $V_{H}C_{H}$ and $V_{H}C_{M}$ (attribution to climate change), and $V_{H}C_{H}$ and $V_{M}C_{H}$ (attribution to land-use change) distributions are significantly different.   If both changes are significant and of the same sign, then the shift is compounding.  If both are significant and in opposite directions then the shift is counteracting.  If either or both changes are not significant, then the change is defined as confounding.

All analysis uses R [@Team2014], RMarkdown [@rmarkdown2016] and R packages `colorspace`, `captioner`, `raster`, `ggplot2`, `mgcv`, `reshape2`, `plyr`, `gridExtra`, `rgdal`, `RColorBrewer`, `analogue`, and `maptools`. All code is hosted on GitHub (http://github.com/PalEON-Project/Composition_Climate). Maps use the public domain Natural Earth Data (http://naturalearthdata.com) products for political and landscape features.

# Results

Regionally, $T_{min}$ increased `r round(clim_change[4], 1)`^o^C since the 1895-1924 normal (`r fig_nums("land_use_clim", display = "cite")`), $T_{max}$ changed by `r round(clim_change[2], 1)`^o^C, while $T_{diff}$ changed by `r round(clim_change[3], 1)`^o^C. Regionally, $P_{ann}$ has increased by `r round(clim_change[1], 1)`mm.  The Palmer Drought Severity Index is closely coherent with $P_{ann}$ trends from 1895 onward and indicates no major climatic trends between 1800 and 1900 (`r fig_nums("land_use_clim", display = "cite")`e), suggesting that early 20^th^ century climates are a reasonable proxy for mid-19^th^ century climates.  When long term records [@burnette2010daily; @baker1985minnesota] are considered, $\Delta t_{min}$ increases to approximately `r round(clim_change[4], 1) + 0.5`^o^C, $\Delta t_{max}$ to `r round(clim_change[2], 1) + 0.35`^o^C and $\Delta t_{diff}$ to `r round(clim_change[2], 1) - 0.15`^o^C.

```{r loss_gain_fig, warning=FALSE, message=FALSE, echo=FALSE, dpi=150}

source("R/figures/gain_loss_plot.R")

#  This takes a bit of time.
loss_melt <- loss_plot()

loss_table <- as.data.frame.matrix(table(loss_melt[[1]]$variable,
                                         loss_melt[[1]]$value))
loss_table <- data.frame(loss_table, 
                         Modern = round((loss_table$Gain + loss_table$Presence) * .064, 0),
                         Historical = round((loss_table$Loss + loss_table$Presence ) * .064, 0))

loss_table[,1:3] <- round(analogue::tran(loss_table[,1:3], method = "proportion") * 100,1)

ggplot2::ggsave(plot = loss_melt[[2]], filename = "figures/Figure_panelloss.tiff",
       dpi = 300, 
       width = 6, 
       height = 3)
```

![](figures/Figure_panelloss.tiff)

`r fig_nums("gain_loss", "Maps showing areas of gain, loss, and continuous presence of tree genera. 'Loss' indicates 8x8km grid cells where PLS data indicated presence of a genus but FIA data do not record presence (light blue). 'Gain' indicates areas where a genus is absent in the PLS data but present in the FIA data (red). 'Continuous presence' represents locations where both FIA and PLS data indicate presence (dark gray). Losses are more common than gains, particularly in the southwestern portion of tree ranges.  Few taxa show gains, although *Populus*, Fraxinus and Tilia show gains of over 10% (Supplemental Material).")`

```{r echo = FALSE}
tab_nums("tab_gain_loss", "Gain and loss (%) of the analyzed tree genera since the pre-settlement era, based on presence or absence within each 8x8km grid cell. Genera are grouped into broadleaved and needleleaved taxa, then ordered alphabetically.  The total regional coverage is 427,000 km^2^ for the PLS data and 331,000 km^2^ for the FIA data. Modern and historical occurence are reported, followed by percentages relative to the total area occupied by the taxon in the pre-settlement and modern eras. Modern and Historical indicate represent range area (100s of km^2^) occupied by the tree taxon for the pre-settlement or modern era.  Taxa are sorted by the percent area lost (Column 3).", display = FALSE)

```

For most tree genera, historical range losses dominate over range gains and are spatially structured (Table 1, Figure 3). Range losses in the upper Midwest are concentrated in the south and west, along the prairie-savanna-forest ecotone (Figure 3). For *Ulmus* and *Larix*, however, losses occur throughout the range.  Losses are highest for *Ostrya*/*Carpinus* and *Larix* when assessed as a proportion of total range (Supplemental Table 1), and range losses are high generally, with an average of `r round(mean(loss_table$Loss), 1)`% across all taxa, *Larix*, *Ulmus* and *Betula* have the largest absolute losses in range within the Upper Midwest.  Gains, conversely, account for no more than `r round(max(loss_table$Gain),1)`% of any taxon's historical range, with the highest gains observed for *Populus* (`r fig_nums("gain_loss", display="cite")`) and averages `r round(mean(loss_table$Gain),1)`% across all taxa.   Spatial patterns of range gains vary widely among tree genera, and often show less spatial bias than losses (Figure 3).  Range gains for *Picea*, *Larix*, *Abies*, *Tsuga*, *Fagus*, *Ostrya*/*Carpinus*, *Tilia*, and *Fraxinus*, for example, all show a dispersed pattern of gains in a relatively few and widely scattered cells, suggesting that these apparent range gains are mainly caused by sampling uncertainty inherent in local-scale plots and stand heterogeneity, rather than systematic range gains.  Conversely, gains for *Populus* and *Acer*, are concentrated in northern Minnesota, Wisconsin, and Michigan, suggesting range infilling, while *Quercus* gains are consistent with some range extension northwards.  *Pinus*, *Juniperus*/*Thuja*, and *Ulmus* range gains are limited but primarily in the southern portion of their historical range, or beyond it.

`r tab_nums("tab_gain_loss", display = "full")`

```{r table1, results="asis", echo=FALSE, eval=TRUE}

sci_names <- c("Larix", "Pinus", "Picea", "Abies", "Tsuga",
               "Thuja/Juniperus", "Populus", "Acer", "Betula", "Fagus", 
               "Ostrya/Carpinus", "Tilia", "Fraxinus", "Ulmus", "Quercus")

rownames(loss_table) <- paste0("*", sci_names, "*")

pander::pandoc.table(loss_table, justify = "left", style = "rmarkdown",
                     emphasize.rownames = FALSE)
```

```{r densities, warning=FALSE, message=FALSE, echo=FALSE, dpi=150}
source("R/resampling_climate.R")
vegclim_table <- resample_climate(climate, agg_dens, pls_data)

source("R/figures/climate_boxplots.R")
boxplot_clim <- clim_boxplots(clim_table = vegclim_table, 
                              filename = 'boxplot_clim')

# Do it with the correction for early historical:
newveg <- vegclim_table
topveg <- vegclim_table
allveg <- vegclim_table

newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmax"] <- newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmax"] - 0.35
newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmin"] <- newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmin"] - 0.5
newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tdiff"] <- newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tdiff"] + 0.15

topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmax"] <- topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmax"] + 0.84
topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmin"] <- topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmin"] - 0.84
topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tdiff"] <- topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tdiff"] + 1.15

allveg$clim[allveg$c.ref == "FIA"] <- topveg$clim[topveg$c.ref == "FIA"]
allveg$clim[allveg$c.ref == "PLSS"] <- newveg$clim[newveg$c.ref == "PLSS"]

boxplot_clim <- clim_boxplots(clim_table = allveg,
                              filename = "boxplot_corrected_all")

boxplot_clim <- clim_boxplots(clim_table = newveg,
                              filename = "boxplot_corrected_pls")

boxplot_clim <- clim_boxplots(clim_table = topveg,
                              filename = "boxplot_corrected_fia")

boxplot_clim <- clim_boxplots(clim_table = vegclim_table,
                              filename = "boxplot_original")

```

![](figures/boxplot_clim.tiff)

`r fig_nums("boxplots", "Changes in the realized climate distributions for tree genera in the upper Midwestern United States. In the box-and-whisker plots, the central line indicates the median, box margins indicate 25^th^ and 75^th^ percentiles, whisker tips indicate 5^th^ and 95^th^ percentiles, and dots indicate outliers beyond the 95^th^ percentile. These shifts in realized climate distributions are produced by the joint effects of historical climate change, pathogen-induced tree mortality, and land-use change.")`

The joint effects of historical climate change and land use on tree-climate distributions are illustrated for *Larix* (Figure 2).  For *Larix* we see clear shifts in climate space when vegetation is held constant (from solid to dashed lines), regardless of whether modern or pre-settlement vegetation distributions are used, indicating that historical climate change plays a strong role in shifting the distribution of *Larix* in climate space across all climate variables.  Historical land use contributes to shifting distributions, particularly for $T_{max}$, where losses at the highest $T_{max}$ values are visible from pre-settlement to modern (Fig. 2d). However, for *Larix*, land use overall appears to exert a weaker influence than climate change.

Shifts in tree-climate distributions are significant for all genera and all climate variables (`r fig_nums("boxplot", display = "cite")`), even accounting for multiple comparisons (all p < 0.0001).  The directionality of shifts between historical and modern climate distributions is consistent across genera: *e.g.*, modern $T_{max}$ is lower for all genera, while $T_{min}$ is higher (`r fig_nums("boxplot", display = "cite")`).  Across all genera, modern $P_{ann}$ is higher than in the past. *Ulmus* shows the largest increase in minimum temperature (`r fig_nums("boxplot", display = "cite")`; $\delta_{Tmin}$ = 4.2 ^o^C), while the smallest increase occurs for *Larix*. Maximum annual temperature declines most for *Ostrya*/*Carpinus* ($\delta_{Tmax}$ = -1.3 ^o^C) and declines least for *Ulmus* ($\delta_{Tmax}$ = -0.4 ^o^C).  *Ulmus* shows the largest increase in mean precipitation ($\delta_{Pann}$ = 86mm), while the smallest change is for *Larix* ($\delta_{Pann}$ = 42mm).

```{r panel_densities, warning=FALSE, message=FALSE, echo=FALSE}
source("R/figures/FigureFour_Panels.R")
```

```{r climate_shifts, warning=FALSE, message=FALSE, echo=FALSE}

source("R/hellinger_dists.R")

#  Returns a list with the ggplot (object 1) and all.hellinger.
hell_plot <- hellinger_plot()

```
![](figures/hellingerplot.tiff)

`r fig_nums("hellPlot", "The total shift in climatic distribution over the last 100 to 150 years for each tree genus ($d_{tot}$) plotted against the amount of shift caused by historical climate change vs. changes in tree distributions ($d_{c} - d_{v}$), the latter mainly due to historical land-use and pathogen outbreaks.  Results are plotted separately for each of the four climate variables considered here. Shifts in distributions are calculated as the Hellinger distance between tree-climate distributions for different combinations of pre-settlement and modern vegetation and climate (Fig. 2).  For $d_{c}-d_{v}$ (Y-axis), negative values indicate a stronger influence of historical land-use on shifting climate distributions, while positive values indicate a stronger influence of historical climate change on shifting distributions.")`

Both the total shift in realized tree-climate distributions and the attribution of these changes to historical land use or climate change varies strongly among taxa and among climate variables (`r fig_nums("hellPlot", display = "cite")`). Shifts in realized climate distributions over the last 100 to 150 years are largest for *Ulmus* ($T_{min}$, $T_{max}$), *Larix* ($T_{max}$), *Tsuga* ($T_{max}$ and $P_{ann}$), and *Fagus* ($P_{ann}$).   Attribution analyses based on $d$ (`r fig_nums("hellPlot", display = "cite")`) reveal that while the imprint of historical land use on shifting climatic niches is important for many taxa, this shift is not uniform across taxa. The climatic signal is strongest for $P_{ann}$, with regional changes in $P_{ann}$ being largely responsible for shifting climate niches, while only *Populus* and *Picea* have seen climate niche shifts due to land-use that are larger than climate shifts. The greatest land use signal across taxa is found in $T_{max}$ (`r fig_nums("hellPlot", display = "cite")`), with effects particularly strong for *Larix*, *Ulmus*, and *Fagus* (all taxa lie below the equality line for $d_{c} - d_{v}$). The large shift in the distribution of *Tsuga* relative to $T_{max}$, conversely, appears to be mainly attributable to historical climate change. For $T_{min}$ (`r fig_nums("hellPlot", display = "cite")`), shifting distributions of *Ulmus* and *Thuja*/*Juniperus* can be mainly attributed to land use change, while *Larix* and *Betula* have large shifts that are equally attributable to historical climate change and land use, and the niche shifts for *Quercus*, *Fraxinus*, *Picea*, *Acer*, and *Tilia* are attributable to changing climate.

`r tab_nums("compTable", "Compounding ($\\Box$), confounding (-) and counteracting ($\\bigcirc$) relationships between land-use and climate change.  When the land-use/climate interaction was regressed against spatial vaariables (latitude, longitude) no significant relationship was found.")`

```{r tests, warning=FALSE, message=FALSE, echo=FALSE, results="asis", eval=TRUE}
source("R/confounding_table.R")

rownames(conf_table) <- NULL

pander::pandoc.table(conf_table, justify = "left", style = "rmarkdown")
knitr::kable(conf_table_corr)
```

The majority of climate - land-use interactions are compounding (n = 30) or confounding (n = 28), while few (n = 2) show counteracting effects (`r tab_nums("compTable", display = "cite")`).  Compounding interactions, where both climate and land use shifts push species distributions in the same direction along the climate gradient, are particularly strong for $T_{max}$ (`r tab_nums("compTable", display = "cite")`), and are consistent with the observed trend of cooling $T_{max}$ (`r fig_nums("land_use_clim", display = "cite")`, `r fig_nums("boxplots", display = "cite")`) and losses concentrated in the southern portions of most species ranges (`r fig_nums("gain_loss", display = "cite")`); both tend to shift tree-climate distributions to cooler portions of climate space. Compounding interactions are also strong for $P_{ann}$ and are caused by both land use and climate change shifting species to a wetter portion of climate space than they occupied historically.  These analyses are insensitive to the type of correction applied to the *p*-value, although running the analysis without a correction factor reduces the counts of insignificant (confounding) relationships and increases the number of  counteracting and compounding interactions.  Patterns of interaction, by species, for $P_{ann}$ and $T_{diff}$ are significantly correlated (${\chi^{2}}_{4}$ = 15.22), but are not significantly correlated for other climate variables.  Analysis included in the supplemental material indicates that a shift to an earlier baseline (using the estimated change rate from [@burnette2010daily] in Manhattan, Kansas) results in no change in this pattern (see "R/confounding_table.R" in S1).

Counteracting interactions, where climate shifts and land use shifts show opposite signs relative to baseline climate, effectively narrowing the realized climatic niche from both directions, exist only for *Ulmus* ($T_{max}$) and *Larix* ($T_{min}$).

Several genera show common patterns in their relationships to climate/land-use interactions.  *Populus*, *Picea*, *Quercus*, *Acer* and *Betula* all show compounding relationships for both $P_{ann}$ and $T_{max}$, either compounding or confounding interactions for $T_{diff}$. and confounding interactions for $T_{min}$ (`r tab_nums("compTable", display = "cite")`).  These taxa all show significant losses in range along the southern and western margins of their distribution (`r fig_nums("gain_loss", display="cite")`), with few gains in those regions.

*Fraxinus*, *Tsuga* and *Tilia* all show confounding interactions with $P_{ann}$, $T_{diff}$ and $T_{min}$, with compounding interactions with $T_{max}$ (`r tab_nums("compTable", display = "cite")`).  Each of these taxa experiences range losses, but patterns of gain and loss differ among taxa.

# Discussion

This work reinforces the importance of historical datasets for examining relationships between species and climate, and their importance for forecasting future species distributions [@moritz2013future;@maguire2015modeling]. Modern surveys sample only a subset of historical distributions, both spatially and within multidimensional climate space.  Our analyses demonstrate that species-climate relationships are changing as a result of land use and climate change over the past few centuries, and adds to our understanding of the interacting effects of climate change and land-use [@pyke2004habitat]. Land use change reduces the correlational structure between species and climate [@devineau2011arid] and is likely to have one of the strongest effects on changes in global biodiversity over the next century [@sala2000global]. While rates of land-use change may be declining in North America and in the upper Midwest [@dale1997landuse;@rhemtulla2009legacies] the legacies of land-use are likely to persist, both in terms of forest composition [@goring2016preset;@kujawa2016theeffect], and the correlational structure between species and climate.

Previous studies have shown that the realized climate niches of tree species shifted during the climate changes accompanying the last deglaciation.  To a first order, many species accommodated past climate changes by shifting their ranges; *i.e.* they maintained a relatively stable distribution in environmental space by shifting their distributions within geographic space.  Range shifts are well documented by paleodata and generally well simulated by species distribution and community-level models driven by paleoclimatic simulations [@martinez2006conservatism;@maguire2015modeling;@prentice1991vegetation].  However, species tracking of past climatic change was imperfect and other studies indicates shifts in species realized niches over time [@maiorano2013building;@pearman2008niche].  Mechanisms for these realized niche shifts include species lagging rapid climate change [@ordonez2013realized], occupying non-analogue climates [@veloz2012no], or shifts in competition resulting from changing patterns of co-occurrence through time [@maiorano2013building].  This study adds to prior work by showing that the land use changes of the last several centuries can also significantly modify species-climate relationships.

In this study, shifts in realized climate niches and the underlying interactions between climate change and land use emerge from three basic processes.  First, climates have changed over the 20^th^ and early 21^st^ centuries, with rising winter minimum temperatures, declining summer temperatures, and higher precipitation.  These climate changes are produced by a combination of natural variability and anthropogenic change; the latter signal is strongest in the second half of the 20^th^ century and early 21^st^ century [@estrada2013statistically]. Second, tree species distributions have not yet fully adjusted to these climate changes, causing tree distributions to be in partial disequilibrium with climate [@svenning2013disequilibrium;@williamsinpress]. Third, anthropogenic land use has selectively eliminated portions of geographic and environmental space from species' potential ranges. Agricultural conversion has largely eliminated open forests in warm and dry regions, while summer temperatures have declined and precipitation increased, resulting in a compounding effect; species occupy cooler habitats with higher precipitation than during historical times. Intensification of land-use in the northern prairie-forest margin shifts species further than expected along the precipitation gradient. While fire and logging have played a role in changing distribution patterns in the north and northeast, they have not resulted in the exclusion or extirpation of taxa from the northern climate space, with the exception of *Tsuga*.

In addition to land use changes, the near total regional collapse of *Ulmus* distributions provides another case in which human agency (in this case, accidentally introduced Dutch elm disease) has affected distributions through near extirpation [@barnes1979stem], although reintroduction efforts are underway. *Castanea dentata* was also effectively wiped out by chestnut blight, with reintroduction efforts underway, and *Tsuga canadensis* and *Fraxinus* spp. are now experiencing rapid mortality due to exotic pests. These pathogen-induced changes in tree distributions do not align neatly with the land use/climate attribution index described here, but they do fall under a broader set of vegetation changes caused by anthropogenic changesfactors, both inadvertent and intentional.

Some of these apparent shifts in climate niches may be due to differences in sampling design rather than to actual shifts in species distributions.  In particular, the FIA dataset may be underrepresenting tree distributions in unforested or semi-forested regions.  If so, estimates of forest loss (`r fig_nums('gain_loss', display='cite')`) and niche shifts (Figures 4 & 5) may be overestimated. Regardless of cause, the differences between the PLS and FIA tree distributions are important for distributional modeling because FIA data provides the most detailed and most widely used source of tree distributional data over the entire contiguous United States [@wang2016landscape;@nieto2015close;@bell2014early].  Hence, any incompleteness in the representation of realized climate niches based on FIA data that are affecting this study should also affect ecological assessments that are based on FIA data.

Anthropogenic agency may also be extending the range of some tree species, for example, as a result of irrigation or landscaping care within urban or rural locations that might otherwise not support forest cover [@brazel2000tale;@groffman2014ecological].  Human agency may also support the extension of forest ecosystems at historical range margins, or preserve historical range margins from land use conversion to agriculture as windrows, or isolated woodlots. Detailed sampling of tree species, particularly in urban and agricultural environments beyond the current expected range limits of species, may provide better representation of species' fundamental climate niches or tolerance niches [@sax2013niche].  However, usage of species distributional data from highly managed ecosystems must account for the intensive management that may be required to maintain species beyond their realized niche.

Compounding and confounding interactions between historical climate change and land use, and the consequent shifts in realized climate space have important implications for the management and genetics of forest communities.  On the one hand, the observation that historical realized niches are broader than modern niches suggests that some tree genera may have more resilience to climate change than expected based on contemporary observations alone. Regionally, annual temperature is expected to increase in Wisconsin by 3.6^o^C in 2055 from a 1980s baseline [@wicci2011climate] and $P_{ann}$ projections range from -17 to 110mm by 2055 relative to 1980s baselines, based on fourteen global climate models from the Climate Model Intercomparison Project [@wicci2011climate]. Every tree genera examined in this study had historical $T_{max}$ values from 0.4 to 1.3^o^C higher than modern estimates, and $P_{ann}$ from 42 to 86mm lower than modern estimates. These higher temperatures and lower precipitation across taxa distributions during the pre-settlement period indicate the potential for greater resilience to climate change than might be assessed using modern data only.  This evidence for historical resilience to warmer and dryer climates may be particularly critical for species that reach their northern or western range limits in the upper Midwest (e.g., *Tsuga*, *Fagus*).

However, projected temperature rises by 2100 far exceed the shifts due to historical land use and climate change.  In addition, coupled migration and adaptation in response to future climate change is likely to be impeded by the extirpation of adapted populations due to land use change and loss of genetic diversity. Local adaptation is an important component of tree responses to climate change [@aitken2008adaptation]. Hence, the loss of genotypes adapted to warmer and drier conditions (the trailing edge of distributions for many taxa in the Upper Midwest) caused by land use may mean more rapid losses at the trailing edges of the species range.  Long distance gene flow of adaptive traits to central and leading populations from the trailing edge [@hu2006seed] is lost when land use conversion to agricultural production results in the extirpation of locally adapted populations.  Thus net gene flow in populations along the remaining trailing edge will be from the central and leading populations, individuals maladapted to warmer conditions. This maladaptive gene flow could potentially amplify the effects of climate change on populations along the forest-prairie boundary leading to increasing risk of local extinction among natural populations [@kremer2012long].

Species distribution models fall along a continuum from more correlative to more process-based [@dormann2012correlation], and correlational species-distribution models are likely to be the most affected by the use of contemporary observational datasets that are biased by historical climate and land-use change and climatic disequilibrium. However, even process based models are partially parameterized from contemporary distributions, particularly with respect to the environmental tolerances of plant functional types [*e.g.* @wullschleger2014plant]. In this case, both correlative and process based models might be affected by the shifting climatic niches demonstrated here, and the underlying processes of spatially biased land-use conversion, changing climates, and climatic disequilibrium.

In summary, historical datasets and multitemporal studies such as this oneadd to our ability to understand the stability and dynamic of species-climate distributions and their adaptive potential in the face of climate change.  The strong changes in climate and land use over the past two centuries, combined with the likelihood that tree distributions are lagging recent climate changes [@svenning2013disequilibrium], create the strong likelihood that tree-climate distributions have shifted and that current tree-climate distributions are not representative of pre-settlement conditions.  Given the importance of modern distributional datasets for predicting species range shifts, compounding interactions between historical climate and land use change may result in the impression of narrower climatic tolerances than actually exist, leading to underestimates of species resilience. Both empirical and process-based vegetation models need to combine, whenever possible, both contemporary and historical information on tree distributions, tree-climate relationships, and their respective shifts through time.

Acknowledgements
---------------------
The authors would like to thanks the large number of individuals who have worked to first, undertake the Public Lands System Survey, to bring the original survey data together, to digitize and standardize much of the survey results, and finally, to assist in interpreting and compiling the data in its present form. In particular, we thank David Mladenoff, Charlie Cogbill, Ed Schools, and the PalEON Settlement Vegetation working group. The authors than Andria Dawson and PalEON participants for helpful comments in the drafting of this manuscript. This material is based upon work carried out by the PalEON Project (http://paleonproject.org), under NSF Award Numbers 1241868 and 1065656.  Maps were made using data from Natural Earth (http://naturalearth.com).

References
---------------