### 2016-12-02 Composition_Ecology.Rmd
Finished read-through of main text, pushing edits

---
bibliography: goringetal_references.bib
csl: ecology.csl
output:
  word_document:
    reference_docx: word_template.docx  
  pdf_document:
    pandoc_args: -V geometry:vmargin=1in -V geometry:hmargin=1in
  md_document:
    variant: markdown_github
pandoc_args: -H margins.sty
---

<!--

```{r setup-knitr, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = " ",
  error = TRUE,
  eval = TRUE,
  cache = TRUE,
  tidy = TRUE

)

library("pander")
library("plyr")

panderOptions('missing', '--')
panderOptions('digits', 4)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.continues', NULL)
pander::panderOptions('table.style', 'rmarkdown')
pander::panderOptions('table.split.table', 300)
grDevices::ps.options(fonts = 'serif')

```
-->

# Effect of historical land-use and climate change on tree-climate relationships in the upper Midwestern United States

**Statement of Authorship**

* Simon J Goring: Co-conceived paper, performed analysis, developed figures and co-wrote the paper
* John W Williams: Co-conceived paper and co-wrote the paper

**Data Accessibility**

* All data for this manuscript, along with code and version-control are available at http://github.com/SimonGoring/CompositionClimate.  In addition, an archived copy of the submitted version is available with the following DOI: XXXX

**Authors**

* Simon J Goring, Department of Geography, University of Wisconsin, Madison, 550 N Park St, Madison WI 53706
* John W Williams, Department of Geography, University of Wisconsin, Madison, 550 N Park St, Madison WI 53706; Center for Climatic Research, University of Wisconsin, Madison, 1225 W Dayton St, Madison WI 53706

**Keywords**

* Anthropocene, climate change, climate disequilibrium, forest inventory and analysis (FIA), fundamental niche, historical ecology, land-use, niche shift, Public Land Survey System, realized niche

**Running Title**

* Causes of recent climate shifts in trees

**Article Type**: Letters

**Number of Words**

* Main Text: 4900
* Text boxes: 0

**Number of References**: 77

**Number of Figures, Tables & Text Boxes**: 6 (5 figures, 1 table)

**Corresponding Author**

* Simon J Goring
  * Department of Geography, University of Wisconsin, Madison, 550 N Park St, Madison WI 53706, goring@wisc.edu

# Abstract
<!-- removed reference to 'higher Tmax' given results of sensitivity analyses.  Also took away the statement about 'reinforces shift to cooler conditions, again given ' 146 words after rewording-->
Contemporary forest inventory data are widely used to understand environmental controls on tree species distributions and to construct models to project forest responses to climate change, but the stability and representativeness of contemporary tree-climate relationships are poorly understood. We show that tree-climate relationships for 15 tree genera in the upper Midwestern US have significantly altered over the last two centuries due to historical land-use and climate change. Realized niches have shifted towards higher minimum temperatures and higher rainfall. A new attribution method implicates both historical climate change and land-use in these shifts, with the relative importance varying among genera and climate variables. Most climate/land-use interactions are compounding, in which historical land-use reinforces  shifts in species-climate relationships toward wetter distributions, or confounding, in which land use complicates shifts towards warmer distributions. Compounding interactions imply that contemporary-based models of species distributions may underestimate species resilience to climate change.

/pagebreak

```{r, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Load climate data and generate the pre-sett and modern bounds:
#  Pre-settlement climate: 1900 - 1930
#  Modern climate: 1960 - 1990

source("R/Setup_R.R")
source("R/load_climate.R") # gives `climate` with elements `then`, `now` and `names`
source("R/load_veg.R")

best.taxa <- c("beech", "ironwood", "hemlock", "fir", "cedar.juniper",
              "basswood", "spruce", "elm", "ash", "pine", "oak",
              "tamarack", "maple", "poplar", "birch")

pls_data <- get_pls("0.9-6")
agg_dens <- get_fia()[,colnames(pls_data)]

land_use <- land_use[land_use$cell %in% c(pls_data$cell, agg_dens$cell),]

source("R/natural_earth.R")

natural <- natural_earth()

```

# Introduction

Contemporary species occurrence databases, and, increasingly, paleohistorical datasets are used to map distributions of species and species traits within environmental space, study underlying processes, and prepare for climate change [@lamanna2014functional;@moritz2013future;@maguire2015modeling;@ivory2016niche;@pearman2008niche]. These distributions of species in geographic and environmental space, combined with future climate scenarios, are used to assess species exposure and sensitivity to climate change, identify species and habitats of concern, set conservation priorities, and prepare for climate-driven shifts in habitat suitability [@pellatt2012using;@iverson2013tree;@iverson1998predicting;@anderson2013nichemodels;@guisan2013predicting]. Forest inventories offer rich observational datasets that have been used to assess early signals of range shifts via the distributions of mature trees and seedlings [@monleon2015evidence;@zhu2012failure] and to parameterize models of current and future tree distributions [@iverson2008estimating;@iverson1998predicting].
<!-- deleted Ordonez 2013 reference to shorten references -->
Increasingly, ecologists are turning to the paleohistorical record to test the assumptions underlying ecological models and improve their predictive capacity [@moritz2013future;@maguire2015modeling]. Projections relying only on contemporary observational data may overlook plasticity in species-climate responses [@maiorano2013building], incompletely delimit species fundamental niches [@jackson2000responses], miss responses to past no-analog climates [@veloz2012no], or fail to detect effects of anthropogenic land-use, and other historical legacies [@svenning2015influence]. Paleoecological records show that forest responses lag abrupt climate change by a few years to centuries [@williamsinpress;@webb1986vegetation;@svenning2015influence], suggesting that tree distributions have yet to fully respond to recent climate changes. In principle, mechanistic dynamic vegetation models [*e.g.*, @leiblein2016modelling] can represent the physiological, demographic, and dispersal processes that cause tree populations to lag climate change.  In practice, however, the parameterization of these models is poorly constrained for slow ecological processes [@moorcroft2006close;@matthes2016benchmarking].

Historical land-use has substantially affected species distributions and vegetation structure [*e.g.*, @gehrig2007tree] and may cause species distributions to incompletely fill available climates, or otherwise be at disequilibrium [@early2014climatic;@sarmento2013impacts;@ivory2016niche]. Anthropogenic land-use is structured and is often biased within climate space, hence, species distributions and protected natural areas often represent a biased selection from the available environmental space [@scott2001nature]. Biasing can confound studies that use contemporary distributional data to understand the environmental controls on species distributions and predict species responses to future climate scenarios [@pyke2004habitat].

Interactions between climate and land-use change can be classified into three types: compounding, confounding and counteracting [modified from @pyke2004habitat]. Compounding effects occur when climatic and land-use changes cause species distributions to shift in the same direction along an environmental gradient.  For example, warming has been compounded by land-use conversion at the lowest elevations in the Sierra Nevada Mountains of California, both causing upslope distributional shifts for butterfly species [@forister2010compounded]. Confounding effects may cause little overall change in species distributions, with land-use pressure occurring across the species' environmental gradient. For example, land-use conversion has excluded Garry oak (*Quercus garryana*) from the deep-soil sites it formerly occupied [@pellatt2014environmental], but land-use has not systematically shifted Garry oak distributions in climate space.  These confounding effects may hinder detection of climatic forcing of historical species distributions. Counteracting effects occur when climate and land-use change act in opposite directions, narrowing the available geographic and climatic space available for species. For example, bird species in the Sierra Nevadas are expected to move downslope due to increasing precipitation [@tingley2012push], but this downslope migration is limited by land-use conversion in the lowlands.

```{r figure1, echo = FALSE, warning = FALSE, message=FALSE, fig.width=6, fig.height=5}

land_use$class[land_use$class == "Water"] <- NA

ext <- c(-100000, 1050000, 600000, 1600000)

land_use_plot <- natural[[2]] +
  geom_tile(data = land_use, aes(x = x, y = y, fill = class)) +
  coord_equal(xlim = ext[1:2],
              ylim = ext[3:4]) +
  scale_fill_manual(values = c("#ff7f00", "#e41a1c", "#33a02c")) +
  theme(legend.position = "none")

source("R/figures/climate_ts.R")

climate_values <- climate_ts(unit.raster, pls_data, climate$now, climate$then)

clim_change <- climate_values$modern$value - climate_values$past$value

ggsave(plot = arrangeGrob(climate_values$plot, land_use_plot, nrow = 1, widths = c(3, 4)),
       filename = "figures/figure_one.tiff")

```
![](Final_Figures/figure_one.tiff)

`r fig_nums("land_use_clim", "*(a-e) Climate change in the upper Midwest over the last two centuries using PRISM data and North American Drought Atlas PDSI reconstructions with a LOESS smooth superimposed for illustration. Modern climate (normals indicated by rectangles).  PDSI shows strong coherence with $P_{ann}$ in this region. (f) Patterns of land-use change in the upper Midwestern United States, with the 'Tension Zone', differentiating sub-boreal from southern broadleafed forests superimposed. land-use data from the NLCD [@jin2013comprehensive].*")`
<!-- WICCI reference below:  Fix author in bibtex to WICCI, i.e. (WICCI, 2011) -->
Tree distributions and community composition in the upper Midwestern US have been heavily influenced by land-use conversion associated with Euro-American Settlement in the 19^th^ and 20^th^ Centuries [@goring2016preset;@schulte2007homogenization;@hanberry2012comparison;@rhemtulla2009historical], outbreaks of exotic pathogens [@barnes1979stem], climate changes accompanying the Little Ice Age [@hotchkiss2007response] and current anthropogenic warming [@wicci2011climate]. Modern forests show greater homogeneity, weaker ecotones, and significantly different species associations than forests surveyed as part of the Public Land Survey System (PLS) prior to widespread land-use conversion in the late 1800s [@goring2016preset;@schulte2007homogenization]. Rhemtulla et al. [-@rhemtulla2009historical] estimate that *Pinus strobus* and *Tsuga canadensis* now occupy only 4% of their original coverage in Wisconsin. Wildland ecosystems have been converted to agricultural and urban land-use [@rhemtulla2009historical], particularly in historical prairie and savanna ecosystems (`r fig_nums("land_use_clim", display = "cite")`f), where almost 50% of southern deciduous savanna was converted to cropland [@rhemtulla2007regional].  In northern forests, ranges of early successional species have expanded and abundances of late-successional species have been reduced due to widespread logging in the 19^th^ and 20^th^ centuries, and ongoing land management [@schulte2007homogenization;@goring2016preset].  Additionally, the formerly dominant *Ulmus americana* has experienced rangewide population collapses caused by the introduction of exotic pathogens [@barnes1979stem], as did *Castanea dentata*, with other tree species currently threatened (*Tsuga canadensis*, *Fraxinus* spp.) [@barnes1979stem;@schlarbaum1998three;@castello1995pathogens].

Here we demonstrate significant shifts in the climatic distributions of tree genera in the upper Midwestern US over the last two centuries due to historical climate and vegetation change, and we diagnose the relative influence of historical climate change and, land-use changes and pathogen outbreaks (for *Ulmus*) on these shifts.  We use gridded historical and contemporary climate observations (PRISM-LT) and estimates of pre-settlement and modern forest vegetation [@goring2016preset] to map the geographic and climatic distributions for 15 major tree taxa in the upper Midwest.  We construct four observed and hypothetical sets of tree-climate relationships, using a 2x2 factorial of pre-settlement and modern vegetation and climates.  We develop a new method, based on Hellinger distances among tree-climate distributions, to quantify the total change and attribute the relative importance of historical vegetation and climate change. We assess whether climate and vegetation changes are compounding, counteracting or confounding. We use this framework to discuss the potential impacts of historical land-use and climate change on inferences about species resilience to climate change, as predicted by species-distribution models based solely on modern distributional data.

# Materials and Methods

## Historical and Modern Vegetation Data:

Estimates of settlement-era forest composition and tree distribution are based on survey data from the PLS [@white1983history;@schulte2001original;@goring2016preset], aggregated to a 64 km^2^grid across the upper Midwestern United States [@goring2016preset]. We briefly review these methods. PLS data originally were collected on a 1.61 x 1.61 km (1 x 1 mi) grid across the upper Midwest from 1830 to 1910 [@schulte2001original;@white1983history]. At each survey point, surveyors noted the closest two to four trees and recorded their common name, distances from survey points, diameters at breast height, and azimuth to trees. Goring et al. [-@goring2016preset] aggregated this data to an 8x8 km grid with an average of 61 survey points (122 trees) per cell. This process included data cleaning, taxonomic standardization, and the application of spatially varying correction factors [@goring2016preset]. Using this data, we produce presence-absence distributions for each taxon, across the upper Midwest for 15 common tree genera. A few common names, such as "Ironwood" can not be clearly resolved to a single tree genus, and so we combine *Ostrya* and *Carpinus*, and *Thuja* and *Juniperus*. We define 'presence' as at least one recorded tree in any 8x8 km grid cell and use the term 'range' as the extent in geographic or climate space of all cells with recorded presence for a specific taxon. <!-- inserted 'geographic or' because we often refer to range gains and losses in the paper -->

Modern forest data are from the US Forest Inventory and Analysis (FIA) program, which provides detailed inventories of contemporary tree distributions, for all regions with >10% forest cover [@woudenberg2010forest]. The FIA uses a nationally standardized sampling procedure; in the Phase 2/3 system, beginning in 1998, each plot consists of four, 7.32m radius, circular sub-plots which are sampled intensively [@woudenberg2010forest].

Comparisons between historical and modern vegetation must address differences in sampling design between the PLS and FIA datasets [@goring2016preset;@kronenfeld2014validating;@kronenfeld2010influence]. Here, the primary solution for maximizing comparability is via aggregation, both spatially, by aggregating records from both datasets to a common scale (8x8 km), and taxonomically, by aggregating taxa to the level of genera. Nonetheless, differences persist.  The FIA contains fewer plots per 8 km cell, with more trees per plot than the PLS data, and FIA plots are sited within forested or partially forested landscapes. while the PLS sampling is uniform. Scattered trees may be present in regions without permanent or temporary FIA plots, complicating the assignments of 'absence' within the FIA. However, the practice of using FIA data as a surrogate for tree distribution data is widespread [@iverson2013tree;@iverson1998predicting], so these complications are general. In prior comparative analyses of the PLS and FIA data, Goring et al. [-@goring2016preset] tested the effect of differential sampling design on the mapped patterns of novel and disappearing forests, and found little effect.

Given the differences in sampling design, FIA data should be better at detecting species presence at the level of individual plots, while detectability at the level of 8km grid cell (the analysis unit of this study) should be similar between FIA and PLS data or higher in the latter.  Within grid cells, there is a tradeoff between the intensive sampling within FIA plots versus more PLS points per grid cell and more uniform sampling of landscape heterogeneity.  The PLS data may be better at detecting tree presence in low-density regions, due to the variable radius of PLS plots.  Both the FIA and PLS datasets both have good coverage in the upper Midwest, with 5,172 8km grid cells with at least one FIA plot and 7,376 grid cells with at least one PLS point.

In mapping tree distributions in the upper Midwest we consider three classes: "gain," where a tree species was present within a grid cell in the FIA, but absent in the PLS; "loss," where a tree was present in a grid cell during the PLS but absent in the FIA; and "continuous presence," where a tree was present within a grid cell in both FIA and PLS eras.  We present gain and loss estimates for all grid cells (Supplementary Table 1) and for the subset of grid cells with both PLS and FIA observations (Supplementary Table  2). For all species, the range within the upper Midwest incompletely samples species ranges, however, the region contains several major ecotones and the southern and western range limit for many tree species [@curtis1959vegetation] so this region comprises an important boundary in geographic and climate space.

## Historical and Modern Climate Data

We use monthly PRISM LT data for 1895 to 2014 CE [@prism2004] for the historical and modern climate data. The PRISM climate variables are available at 800m resolution and are interpolated from station locations using elevation, aspect and other data.  We resample estimates for mean daily July temperature ($T_{max}$), mean daily January temperature ($T_{min}$), annual temperature range ($T_{diff}$: $T_{max}$ - $T_{min}$), and annually summed daily precipitation ($P_{ann}$) to the 8 x 8km grid used for PLS and FIA data. These climate variables represent three major climatic controls on plant distributions: summer warmth, winter minimum temperatures, and moisture availability; these or similar bioclimatic variables are widely used in plant distributional modeling [@woodward1987climate;@iverson1998predicting;@maiorano2013building].

Three periods were selected for calculation of climate normals: 1895 to 1919 for the historical period and two alternatives for the modern period; 1950 - 1979 representing the average period of establishment for stand in the upper Midwest and 1985 - 2014 covering the period of sampling for the FIA data. We report results for the 1985 - 2014 normals, with the 1950-1980 results shown in Supplementary Information. There is an unavoidable temporal mismatch between the historical climate data and the PLS survey period, on the order of 50 to 70 years; few meteorological records exist in this region prior to 1895 CE, and no gridded dataset exists at high spatiotemporal resolution for this time period for all climate variables.  The effects of temporal mismatch should be minor given that Northern Hemisphere temperature trends were generally small during the 19^th^ century [@mann2009global].  To check whether an 1895 to 1919 CE climate normal is a reasonable surrogate for pre-settlement climates we compare the instrumental record against dendroclimatic reconstructions of the Palmer Drought Severity Index [@heim2002review] (`r fig_nums("land_use_clim", display = "cite")`) and to early temperature observations from US military forts and observer stations, which extend to the 1830s [@baker1985minnesota;@burnette2010daily].

## Shifts in Species Distributions in Climate Space

![](Final_Figures/figure_two.tiff)
<!-- Is Figure 2 for Larix or Quercus?  Legend here says Larix, later in text says Quercus -->
`r fig_nums("fourPanel", "(a) The 2x2 factorial design used to diagnose the effects of historical climate and vegetation change on observed shifts in tree-climate relationships. Observed and hypothetical tree-climate relationships for one of four combinations of pre-settlement vegetation (pink fill, left column) or modern vegetation (blue fill, right column) and pre-settlement climate (solid border, top row) or modern climate (dashed border, lower row). (b-e) Representative normalized probability densities for *Larix* for each of the four combinations of climate and vegetation states, plotted relative to $T_{max}$ (b), $T_{diff}$ (c), $T_{min}$ (d), and $P_{ann}$ (e).")`

Distributions of tree genera within climate space are shown using box plots and described as univariate probability density functions (PDFs) of tree species presence in vegetation ($V$) along a climate ($C$) gradient, estimated using an unweighted Gaussian kernel density estimator, using R's `density` function.  Differences among univariate PDFs are quantified for each climate variable using Hellinger distances [@matthes2016benchmarking] among the univariate PDFs representing the four combined states of vegetation and climate ($V_{H}C_{H}$, $V_{H}C_{M}$, $V_{M}C_{H}$, $V_{M}C_{M}$, where $M$ and $H$ represent modern and historical conditions). The Hellinger distance for two discrete probability distributions is:

$H(P, Q) = \frac{1}{\sqrt{2}} \; \sqrt{\sum_{i=1}^k (\sqrt{p_i} - \sqrt{q_i})^2}$

where $P$ and $Q$ are the distributions, with a common index $i$.

The total observed shift ($d_{tot}$) between a historical and modern PDF is the Hellinger distance between $V_{M}C_{M}$ and $V_{H}C_{H}$ (`r fig_nums("fourPanel", display = "cite")`a: top left to bottom right). The conditions $V_{M}C_{H}$ and $V_{H}C_{M}$ are synthetic, generated by superimposing historical climate patterns on modern vegetation distributions and *vice versa*.  The effect of climate change on shifting PDFs ($d_{c}$) is the difference between PDFs using the early climate normals ($C_{H}$) and modern climate normals ($C_{M}$) overlaid on PLS vegetation (e.g., $V_{H}C_{H}$ versus $V_{H}C_{M}$; `r fig_nums("fourPanel", display = "cite")`a; left column). The effect of vegetation change on shifting PDFs is $d_{v}$, the Hellinger distance between $V_{H}C_{H}$ and $V_{M}C_{H}$ (top row of Figure 2a). Because Euro-American land-use is the dominant driver of vegetation change in the upper Midwest over the past two centuries (Introduction), we often attribute shifts recorded by $d_{v}$ to the land-use change associated with Euro-American settlement.

We develop an attribution index, ($\Delta_{d} = d_{c} - d_{v}$), such that $\Delta_{d} < 0$ indicates stronger influence of historical land-use on shifting climate distributions, while $\Delta_{d}> 0$ indicates a stronger influence of historical climate change.

To assess whether the effects of historical land-use and climatic changes are compounding, counteracting or confounding we perform t-tests, using the conservative Bonferroni correction to account for multiple tests and adjustments for spatial autocorrelation (Supplementary Information).  For each taxon and climate variable pair we test whether $V_{H}C_{H}$ and $V_{H}C_{M}$ (attribution to climate change), and $V_{H}C_{H}$ and $V_{M}C_{H}$ (attribution to vegetation change) distributions are significantly different (Figure 2).   If both changes are significant and of the same sign, then the shift is compounding.  If both are significant and in opposite directions then the shift is counteracting.  If climate change is significant, but land-use is not significant, then the change is defined as confounding.  <!-- We should define the other as confounding as well, since it is, it makes it seem like climate has changed. -->

## Analysis

All analyses use R [@Team2014], RMarkdown [@rmarkdown2016] and R packages `colorspace`, `captioner`, `raster`, `ggplot2`, `mgcv`, `reshape2`, `plyr`, `gridExtra`, `rgdal`, `RColorBrewer`, `analogue`, and `maptools`, `dplyr` and `purrr`. All code is hosted on GitHub (http://github.com/PalEON-Project/Composition_Climate). Maps use the public domain Natural Earth Data (http://naturalearthdata.com) products for political and landscape features.

# Results

Regionally, $T_{min}$ increased `r round(clim_change[4], 1)`^o^C between the 1895-1924 and 1985-2014 normals (`r fig_nums("land_use_clim", display = "cite")`), $T_{max}$ changed by `r round(clim_change[2], 1)`^o^C, while $T_{diff}$ changed by `r round(clim_change[3], 1)`^o^C (see Supplementary Information for estimates based on the 1950 - 1979 normals). $P_{ann}$ increased by `r round(clim_change[1], 1)`mm.  The Palmer Drought Severity Index is correlated to $P_{ann}$ trends from 1895 onward (r = 0.43) and, with LOESS regression, indicates no major climatic trends between 1800 and 1900 (`r fig_nums("land_use_clim", display = "cite")`e), suggesting that early 20^th^-century precipitation is a reasonable proxy for mid-19^th^-century precipitation.  Long term observer and military fort observations suggest warming during the 19^th^ century, with increases in mean annual temperature at Fort Snelling, Minnesota [@baker1985minnesota] and Fort Leavenworth, Kansas [@burnette2010daily]. The latter shows faster increases during winter (`r round(clim_change[4], 1) + 0.5`^o^C) than summer(`r round(clim_change[2], 1) + 0.35`^o^C). Hence, the PRISM-based estimates of historical climate change may be too conservative for $T_{min}$ and $T_{diff}$, and too large for $T_{max}$.

```{r loss_gain_fig, warning=FALSE, message=FALSE, echo=FALSE, dpi=150}

source("R/figures/gain_loss_plot.R")

#  This takes a bit of time.
if ('loss_melt.rds' %in% list.files('data/output/')) {
  loss_melt <- readRDS('data/output/loss_melt.rds')
} else {
  loss_melt <- loss_plot()
  saveRDS(loss_melt, file = 'data/output/loss_melt.rds')
}

loss_table <- as.data.frame.matrix(table(loss_melt[[1]]$variable,
                                         loss_melt[[1]]$value))
loss_table <- data.frame(loss_table,
                         Modern = round((loss_table$Gain + loss_table$Presence) * .064, 0),
                         Historical = round((loss_table$Loss + loss_table$Presence ) * .064, 0))

loss_table[,1:3] <- round(analogue::tran(loss_table[,1:3], method = "proportion") * 100,1)

ggplot2::ggsave(plot = loss_melt[[2]], filename = "figures/Figure_panelloss.tiff",
       dpi = 300,
       width = 6,
       height = 3)
```

![](figures/Figure_panelloss.tiff)

`r fig_nums("gain_loss", "Maps showing areas of gain, loss, and continuous presence of tree genera. 'Loss' indicates 8x8km grid cells where PLS data indicated presence of a genus but FIA data do not record presence (light blue). 'Gain' indicates areas where a genus is absent in the PLS data but present in the FIA data (red). 'Continuous presence' represents locations where both FIA and PLS data indicate presence (dark gray). Losses are more common than gains, particularly in the southwestern portion of tree ranges.  Few taxa show gains, although *Populus*, Fraxinus and Tilia show gains of over 10% (Supplemental Material).")`

All tree genera show higher losses than gains (Supplemental Table 1, Figure 3). Range losses in the upper Midwest are concentrated in the south and west, along the prairie-savanna-forest ecotone (Figure 3). For *Ulmus* and *Larix*, however, losses occur throughout the range.  Losses are highest for *Ostrya*/*Carpinus* and *Larix* when assessed as a proportion of total range (Supplemental Table 1), and range losses are high generally, with an average of `r round(mean(loss_table$Loss), 1)`% across all taxa, *Larix*, *Ulmus* and *Betula* have the largest absolute losses in range within the upper Midwest.  Gains, conversely, account for no more than `r round(max(loss_table$Gain),1)`% of any taxon's historical range, with the highest gains observed for *Populus* (`r fig_nums("gain_loss", display="cite")`) and average `r round(mean(loss_table$Gain),1)`% across all taxa.  Spatial patterns of gains vary widely among tree genera, and often show less spatial bias than losses (Figure 3).  *Picea*, *Larix*, *Abies*, *Tsuga*, *Fagus*, *Ostrya*/*Carpinus*, *Tilia*, and *Fraxinus*, for example, all show a dispersed pattern of gains in few and widely scattered cells, suggesting that these apparent range gains are mainly caused by sampling uncertainty inherent in local-scale plot and stand heterogeneity, rather than systematic range gains.  Conversely, gains for *Populus* and *Acer*, are concentrated in northern Minnesota, Wisconsin, and Michigan, suggesting range infilling, while *Quercus* gains are consistent with some range extension northwards.  *Pinus*, *Juniperus*/*Thuja*, and *Ulmus* range gains are limited but primarily in the southern portion of their historical range, or beyond it. These patterns of range gains and losses are generally insensitive to choice of all gridcells versus the subset of gridcells with data in both time periods, except for *Quercus*, which has a much lower estimate of range loss when only the subset is used (46% vs. 25.7%, Figure 3, Table 1, Supplementary Figure 1, Supplementary Table 2).

```{r densities, warning=FALSE, message=FALSE, echo=FALSE, dpi=150}

sci_names <- c("Larix", "Pinus", "Picea", "Abies", "Tsuga",
               "Thuja/Juniperus", "Populus", "Acer", "Betula", "Fagus",
               "Ostrya/Carpinus", "Tilia", "Fraxinus", "Ulmus", "Quercus")
rownames(loss_table) <- paste0("*", sci_names, "*")

source("R/resampling_climate.R")
vegclim_table <- resample_climate(climate, agg_dens, pls_data)

source("R/figures/climate_boxplots.R")
boxplot_clim <- clim_boxplots(clim_table = vegclim_table,
                              filename = 'boxplot_clim')

# Do it with the correction for early historical:
newveg <- vegclim_table
topveg <- vegclim_table
allveg <- vegclim_table

newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmax"] <- newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmax"] - 0.35
newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmin"] <- newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tmin"] - 0.5
newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tdiff"] <- newveg$clim[newveg$c.ref == "PLSS" & newveg$climate == "tdiff"] + 0.15

topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmax"] <- topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmax"] + 0.84
topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmin"] <- topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tmin"] - 0.84
topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tdiff"] <- topveg$clim[topveg$c.ref == "FIA" & topveg$climate == "tdiff"] + 1.15

allveg$clim[allveg$c.ref == "FIA"] <- topveg$clim[topveg$c.ref == "FIA"]
allveg$clim[allveg$c.ref == "PLSS"] <- newveg$clim[newveg$c.ref == "PLSS"]

boxplot_allveg <- clim_boxplots(clim_table = allveg,
                              filename = "boxplot_corrected_all")

boxplot_plss <- clim_boxplots(clim_table = newveg,
                              filename = "boxplot_corrected_pls")

boxplot_clim <- clim_boxplots(clim_table = topveg,
                              filename = "boxplot_corrected_fia")

boxplot_clim <- clim_boxplots(clim_table = vegclim_table,
                              filename = "boxplot_original")

```

![](figures/boxplot_clim.tiff)

`r fig_nums("boxplots", "Changes in the realized climate distributions for tree genera in the upper Midwestern United States produced by the joint effects of historical climate change and land-use change. The central line indicates the median, box margins indicate 25^th^ and 75^th^ percentiles, whisker tips indicate 5^th^ and 95^th^ percentiles, and dots indicate outliers beyond the 95^th^ percentile.")`

Shifts in tree-climate distributions are significant for all genera and climate variables (`r fig_nums("boxplot", display = "cite")`), even accounting for multiple comparisons and spatial autocorrelation (all p < 0.0001).  The directionality of shifts is consistent across genera: *e.g.*, modern $T_{max}$ is lower for all genera, while $T_{min}$ and $P_{ann}$ are higher (`r fig_nums("boxplot", display = "cite")`).  *Ulmus* shows the largest increase in minimum temperature (`r fig_nums("boxplot", display = "cite")`; $\delta_{Tmin}$ = 4.2 ^o^C), while *Larix* has the smallest. Maximum annual temperature declines most for *Ostrya*/*Carpinus* ($\delta_{Tmax}$ = -1.3 ^o^C) and least for *Ulmus* ($\delta_{Tmax}$ = -0.4 ^o^C).  *Ulmus* shows the largest increase in mean precipitation ($\delta_{Pann}$ = 86mm), and *Larix* the smallest ($\delta_{Pann}$ = 42mm). These patterns are unchanged by sensitivity analysis in which the historical temperature changes are augmented by estimated temperature trends reported from the historical records, but the $T_{max}$ changes are sensitive to choice of modern climate normal (Supplemental Information).

```{r panel_densities, warning=FALSE, message=FALSE, echo=FALSE}
source("R/figures/FigureFour_Panels.R")
```

```{r climate_shifts, warning=FALSE, message=FALSE, echo=FALSE}

source("R/hellinger_dists.R")

#  Returns a list with the ggplot (object 1) and all.hellinger.
hell_plot     <- hellinger_plot()
# For suplemental data:
hell_plot_pls <- hellinger_plot(newveg, 'hellinger_pls')
hell_plot_fia <- hellinger_plot(topveg, 'hellinger_fia')

```
![](figures/hellingerplot.tiff)

`r fig_nums("hellPlot", "The total shift in climatic distribution over the last 100 to 150 years for each tree genus ($d_{tot}$) plotted against the amount of shift caused by historical climate change vs. changes in tree distributions ($d_{c} - d_{v}$), the latter mainly due to historical land-use.  Results are plotted separately for each of the four climate variables considered here. Shifts in distributions are calculated as the Hellinger distance between tree-climate distributions for different combinations of pre-settlement and modern vegetation and climate (Fig. 2).  For $d_{c}-d_{v}$ (Y-axis), negative values indicate a stronger influence of historical land-use on shifting climate distributions, while positive values indicate a stronger influence of historical climate change on shifting distributions.")`
<!-- Does Fig. 2 show Larix or Quercus?  This paragraph states Quercus; legend states Larix -->
Both the total shift in realized tree-climate distributions and the attribution of these changes to historical vegetation or climate change vary among taxa and among climate variables (`r fig_nums("hellPlot", display = "cite")`).  The joint effects of historical climate and vegetation change on tree-climate distributions are illustrated for *Quercus* (`r fig_nums("fourPanel", display = "cite")`).  For *Quercus* we see changes due to both climate (`r fig_nums("fourPanel", display = "cite")`b to d) and due to vegetation change (`r fig_nums("fourPanel", display = "cite")`b to c), but a stronger apparent shift for historical climate change, reflected in the attribution index $\Delta_{d} (`r fig_nums("hellPlot", display = "cite")`a).

<!-- Note that figure numbering is still off - Fig 6 should be Fig 5.  Needs fixing prior to submission-->
Shifts in realized climate distributions over the last 100 to 150 years are largest for *Ulmus* ($T_{min}$, $T_{max}$), *Larix* ($T_{max}$), *Tsuga* ($T_{max}$ and $P_{ann}$), and *Fagus* ($P_{ann}$).   Attribution analyses based on $\Delta_{d} (`r fig_nums("hellPlot", display = "cite")`) reveal that while the imprint of historical vegetation change on shifting climatic niches is important for many taxa, the effect varies across taxa. The climatic signal is strongest for $P_{ann}$, with regional changes in $P_{ann}$ being largely responsible for shifting climate niches, while only *Populus* and *Picea* have seen precipitation niche shifts that are slightly more attributable to land-use than climate shifts (`r fig_nums("hellPlot", display = "cite")`a). The greatest land-use signal across taxa is found in $T_{max}$ (`r fig_nums("hellPlot", display = "cite")`), with effects particularly strong for *Larix*, *Ulmus*, and *Fagus* (all taxa lie below the equality line for $d_{c} - d_{v}$). The large shift in the distribution of *Tsuga* relative to $T_{max}$, conversely, appears to be mainly attributable to historical climate change. For $T_{min}$ (`r fig_nums("hellPlot", display = "cite")`), shifting distributions of *Ulmus* and *Thuja*/*Juniperus* can be mainly attributed to land-use change, while *Larix* and *Betula* have large shifts that are equally attributable to historical climate change and land-use. The $T_{min}$ shifts for *Quercus*, *Fraxinus*, *Picea*, *Acer*, and *Tilia* are attributed to changing climate.

`r tab_nums("compTable", "Compounding ($\\Box$), confounding (-) and counteracting ($\\bigcirc$) relationships between land-use and climate change.  When the land-use/climate interaction was regressed against spatial variables (latitude, longitude) no significant relationship was found.")`

```{r tests, warning=FALSE, message=FALSE, echo=FALSE, results="asis", eval=TRUE}

source("R/confounding_table.R")

rownames(conf_table) <- NULL

colnames(conf_table) <- c("Taxon", "$P_{ann}$", "$T_{max}$", "$T_{diff}$", "$T_{min}$")

conf_table[,1] <- c("Larix", "Pinus", "Picea", "Abies", "Tsuga",
               "Thuja/Juniperus", "Populus", "Acer", "Betula", "Fagus",
               "Ostrya/Carpinus", "Tilia", "Fraxinus", "Ulmus", "Quercus")

knitr::kable(conf_table)

```
<!-- Given new Table 1, have rewritten to state that all climate-veg interaction are compounding or confounding -->
<!-- Last sentence in para may need tweaking, if Supp Table 4 is changed - see comments there. -->
All climate-vegetation interactions are compounding ($\\Box$; n = 30) or confounding ($-$; n = 30), with no counteracting effects found (`r tab_nums("compTable", display = "cite")`). Compounding interactions, where both climate and land-use shifts push species distributions in the same direction along the climate gradient, are strongest for $T_{max}$ (`r tab_nums("compTable", display = "cite")`), and are consistent with the observed trend of cooling $T_{max}$ (`r fig_nums("land_use_clim", display = "cite")`, `r fig_nums("boxplots", display = "cite")`) and losses concentrated in the southern portions of most species ranges (`r fig_nums("gain_loss", display = "cite")`); both tend to shift tree-climate distributions to cooler portions of climate space ($-_c$). Compounding interactions are also strong for $P_{ann}$ and are caused by both land-use and climate change shifting species to a wetter portion of climate space ($+_c$) than they occupied historically.  The patterns of confounding and compounding interactions are insensitive to alternative estimates of 19^th^ century temperature changes (Supplementary Information Table 3).  However, the apparent $T_{max}$ cooling and number of compounding interactions are reduced when 1950 - 1980 climate normals are used instead of 1985-2014 normals (Supplementary Table 4).

Several genera show common patterns in their relationships to climate/land-use interactions.  *Populus*, *Picea*, *Quercus*, *Acer* and *Betula* all show compounding relationships for both $P_{ann}$ and $T_{max}$, either compounding or confounding interactions for $T_{diff}$. and confounding interactions for $T_{min}$ (`r tab_nums("compTable", display = "cite")`).  These taxa all show significant losses in range along the southern and western margins of their distribution (`r fig_nums("gain_loss", display="cite")`), with few gains in those regions.  *Fraxinus*, *Tsuga* and *Tilia* all show confounding interactions for $P_{ann}$, $T_{diff}$ and $T_{min}$, and compounding interactions for $T_{max}$ (`r tab_nums("compTable", display = "cite")`).  Each of these taxa experiences range losses, but patterns of gain and loss differ among taxa.

# Discussion

## Key Findings and Implications

This work reinforces the utility of historical datasets for examining relationships between species and climate, and their importance for forecasting future species distributions [@moritz2013future;@maguire2015modeling]. Modern surveys sample a subset of historical distributions, both spatially and within multidimensional climate space.  Our analyses demonstrate that species-climate relationships are changing as a result of land-use and climate change over the past two centuries, and add to our understanding of the interacting effects of climate change and land-use [@pyke2004habitat]. Land-use change reduces the correlational structure between species and climate [@devineau2011arid] and has been one of the strongest drivers of changes in global biodiversity over the 20th and 21st centuries [@sala2000global]. While rates of land-use change may be declining in North America and in the upper Midwest [@dale1997landuse;@rhemtulla2009legacies], the legacies of land-use are likely to persist, both in terms of forest composition [@goring2016preset;@kujawa2016theeffect], and the correlational structure between species and climate.

Previous studies have shown that the realized climate niches of tree species shifted during the climate changes accompanying the last deglaciation [@pearman2008niche;@maiorano2013building;@veloz2012no].  Many species accommodated past climate changes by shifting their ranges; *i.e.* they maintained a relatively stable distribution in environmental space by shifting their distributions within geographic space.  Range shifts are well documented by paleodata and generally well simulated by species distribution and community-level models driven by paleoclimatic simulations [@martinez2006conservatism;@maguire2015modeling;@prentice1991vegetation].  Mechanisms for realized niche shifts include species lagging rapid climate change [@ordonez2013realized], non-analogue climates [@veloz2012no], or shifts in competition resulting from changing species associations [@maiorano2013building].  This study adds to prior work by showing that the land-use changes of the last several centuries can also significantly modify species-climate relationships.

In this study, shifts in realized climate niches and the underlying interactions between climate change and land-use emerge from three processes.  First, climates have changed over the 20^th^ and early 21^st^ centuries, with rising winter minimum temperatures, higher precipitation, and variable summer temperatures.  These trends are produced by a combination of natural variability and anthropogenic change; the latter signal strengthens beginning in the second half of the 20^th^ century [@estrada2013statistically]. Second, tree species distributions presumably have not yet fully adjusted to these climate changes, causing tree distributions to be in partial disequilibrium with climate [@svenning2013disequilibrium;@williamsinpress]. Third, anthropogenic land-use has selectively eliminated portions of geographic and environmental space from species' potential ranges (Figure 1f). Agricultural conversion has greatly reduced the extent of savanna and open forests in southern Wisconsin and Minnesota, causing tree species to occupy cooler habitats with higher precipitation than historically. While logging and altered fire regimes have altered community composition in the north, they have not widely excluded taxa from the northern regions and associated climates, except for *Tsuga* (`r fig_nums("gain_loss", display="cite")`), a shade-tolerant and late-successional taxon.

In addition to land use, the regional collapse of *Ulmus* distributions points to the importance of forest pathogens as mechanism by which human agency (introduced Dutch elm disease) has affected tree-climate relationships [@barnes1979stem]. *Castanea dentata* was also effectively wiped out by chestnut blight, with reintroduction efforts underway, and *Tsuga canadensis* and *Fraxinus* spp. are now experiencing rapid mortality due to exotic pests [@castello1995pathogens;@schlarbaum1998three]. Hence, land use is not the sole driver of vegetation changes reported here, although it the dominant one.

<!-- Fix the WICCI ref, as above. -->
Compounding and confounding interactions between historical climate change and land-use have important implications for the management and genetics of forest communities. Some tree genera may have more resilience to climate change than expected based on contemporary observations alone. Regionally, annual temperature is expected to increase in Wisconsin by 3.6^o^C in 2055 with $P_{ann}$ projections ranging from -17 to 110mm from a 1980s baseline [@wicci2011climate]. Every tree genera examined in this study had historical $T_{max}$ values from 0.4 to 1.3^o^C higher than modern estimates (primarily attributable to land use change, Figure 5), and $P_{ann}$ from 42 to 86mm lower than modern estimates (primarily attributable to climate change, Figure 5). The higher temperatures and lower precipitation associated with pre-settlement tree distributions may suggest greater resilience to climate change than might be assessed using modern data only.  This evidence for historical resilience to warmer and dryer climates is particularly critical for species that reach their range limits in the upper Midwest (e.g., *Tsuga*, *Fagus*).

Conversely, adaptation to future climate change may be impeded by the historical reduction of locally adapted populations due to past land-use change. Local adaptation is an important component of tree responses to climate change [@aitken2008adaptation], so the historical loss of genotypes adapted to warmer and drier conditions may mean more rapid losses at the trailing edges of the species range.  Similarly, long-distance gene flow of adaptive traits to central and leading populations from the trailing edge [@hu2006seed] may be lost when land-use conversion to agricultural production results in the extirpation of locally adapted populations.  Net gene flow in populations along the remaining populations may be weighted towards individuals maladapted to warmer conditions, which could amplify the effects of climate change at the trailing edge [@kremer2012long].

Correlational species-distribution models are likely to be the most affected by the use of contemporary observational datasets that are biased by historical climate and land-use change and climatic disequilibrium. However, even mechanistic models are partially parameterized from contemporary distributions, particularly with respect to the environmental tolerances of plant functional types [*e.g.* @wullschleger2014plant]. Hence, the predictive ability of both correlative and process based models should be sensitive to the shifting climatic niches demonstrated here, and the underlying processes of spatially biased land-use conversion, changing climates, and climatic disequilibrium.  One solution is to combine historical and contemporary data when building species distribution models [@maiorano2013building].

## Potential Limitations

Several potential limitations are associated with this study: 1) differences in sampling design between FIA and PLS datasets and effects on inferred tree distributions, 2) choice of time period for the historical climate normal, and 3) choice of time period  for the modern climate normal. We address each in turn and explain why the main finding of this paper – that tree-climate relationships have shifted over the last few centuries in response to historic land-use and climate change – is robust.

The first point is the most critical, because some of the reported range losses may be an artifact of incomplete FIA sampling of tree distributions in unforested or semi-forested regions. If so, estimates of forest loss (Figure 3) and niche shifts (Figures 4 & 5) may be overestimated, in particular along the warmest and driest region of the tree distributions in the upper Midwest. However, the range gain/loss/presence estimates are not qualitatively affected by analysis of all gridcells (Figure 3, Table 1) versus the subset of gridcells with both PLS and FIA data (Supplementary Information).  However, it remains possible that FIA plots have a lower detectability than PLS data because of the fewer plots per 8km gridcell.  Regardless of cause, the differences between the PLS and FIA tree distributions are important for distributional modeling because FIA data provides the most detailed and most widely used source of tree distributional data over the entire contiguous United States [@bell2014early;@nieto2015close;@wang2016landscape]. Hence, any incompleteness in the representation of realized climate niches based on FIA data that are affecting this study should also affect ecological assessments based on FIA data.

Our results are generally insensitive to the choice of climatic normal (potential limtitations 2 & 3), except for analyses related to $T_{max}$. The reported 19th-century warming from US forts means that the PRISM-based estimates of changes in $T_{min}$ and $T_{diff}$ are conservative.  Similarly, patterns of change for $T_{min}$, $T_{diff}$ and $P_{ann}$ remain robust to choice of modern climate normal (1950 -– 1980 vs. 1995 -- 2014) (Supplementary Information).  The reported change in $T_{max}$, however, is affected in both cases, suggesting that actual $T_{max}$ trends are small and sensitive to choice of climatic normal. Note that the effects of historic land-use on species-climate relationships are insensitive to choice of climate normal (because climate is held constant when calculating $d_v$), so the effects of land-use on species shifting to cooler portions of $T_{max}$ (Figure 5, Table 1) is not affected by choice of climate normal for $T_{max}$.  Independent analyses using pollen-based climate reconstruction shows significant shifts in pollen-climate relationships following Euro-American settlement in the region [@jacques2008pre], which further indicates that vegetation-climate relationships have changed since settlement.

## Summary

Historical datasets and multitemporal studies such as this one add to our understanding of the stability and dynamic of species-climate distributions and their adaptive potential in the face of climate change.  Land use over the past two centuries and its legacies, combined with the likelihood that tree distributions are lagging recent climate changes, create the strong likelihood that tree-climate distributions in the upper Midwest have shifted and that current tree-climate distributions are not representative of pre-settlement conditions.  The main conclusions of this study are robust to changes in vegetation sampling design over time and choice of historic and modern climatic normals, although both concerns are important and deserve further study.  Given the importance of modern distributional datasets for predicting species range shifts, compounding interactions between historical climate and land-use change may result in the impression of narrower climatic tolerances than actually exist, leading to underestimates of species resilience. Both empirical and process-based vegetation models need to combine, whenever possible, both contemporary and historical information on tree distributions, tree-climate relationships, and their respective shifts through time.

Acknowledgements
---------------------
The authors would like to thank the large number of individuals who have worked to undertake the Public Lands System Survey, to compile the original surveyor notes, to digitize and standardize the survey results, and finally, to assist in interpreting and compiling the data in its present form. In particular, we thank David Mladenoff, Charlie Cogbill, Ed Schools, and the PalEON Settlement Vegetation working group. The authors thank Andria Dawson and PalEON participants for helpful comments in the drafting of this manuscript. This material is based upon work carried out by the PalEON Project (http://paleonproject.org), under NSF Award Numbers 1241868 and 1065656.  Maps were made using data from Natural Earth (http://naturalearth.com).

References
---------------