### 2016-12-04 SupplementaryInformation.Rmd

---
output:
  word_document:
    reference_docx: word_template.docx
---

# Spatial Autocorrelation

We include the effects of autocorrelation in the t-tests using the `SpatialPack` function `modified.ttest`. Unfortunately, the `modified.ttest` only performs a pairwise t-test.  Pairwise t-tests are inappropriate for these analyses, because, for the $V_{H}C_{H}$ - $V_{M}C_{H}$ comparison, only the overlapping points would be included, and it would show zero change since climate is unchanged.

To overcome this challenge, we obtained a modified estimate of degrees of freedom ($df$) that accounts for spatial autocorrelation based on Moran's I from `modified.ttest` for each of the two spatial processes. To then produce a difference of the means that is not pairwise, we obtain the corrected df from `modified.ttest`, and recalculate the p-value for the t-test using the $t$ estimate from the `t.test`. The corresponding code is available in [`confounding_table.R`](https://github.com/PalEON-Project/Composition_Climate/blob/master/R/confounding_table.R).

# Climate changes between 1895-1924 and 1950-1979  (vs. main paper analysis of 1895-1924 and 1990-2014)

If a 1950-1979 climate normal is used instead of 1990-2014 climate normal, then between 1895-1924 and 1950-1979, the regional climate changes are as follows: ($T_{min}$) increased 0.5^o^C between 1895-1924 and 1950-1979, ($T_{max}$) increased by 0.35^o^C, while T_diff changed by -0.15^o^C

# Supplemental Tables

**Supplemental Table 1**:  Gain and loss of the analyzed tree genera since the pre-settlement era. The first three columns (Gain, Loss, Presence) are expressed as percentages relative to the total area occupied by the taxon in the pre-settlement and modern eras. Modern and Historical indicate represent range area (units: 10^2^ km^2^) occupied by the tree taxon for the pre-settlement or modern era. Taxa are sorted by the percent area lost (Column 3).  In this analysis, all PLS and FIA grid cells were used and a taxon was assumed have been gained/lost for a given grid cell if that grid cell had data in one time period but not the other.

```{r, results = 'as-is', warning=FALSE, message=FALSE, echo=FALSE}

loss_melt <- readRDS('data/output/loss_melt.rds')
loss_table <- as.data.frame.matrix(table(loss_melt[[1]]$variable,
                                         loss_melt[[1]]$value))
loss_table <- data.frame(loss_table,
                         Modern = round((loss_table$Gain + loss_table$Presence) * .064, 0),
                         Historical = round((loss_table$Loss + loss_table$Presence ) * .064, 0))

loss_table[,1:3] <- round(analogue::tran(loss_table[,1:3], method = "proportion") * 100,1)

sci_names <- c("Larix", "Pinus", "Picea", "Abies", "Tsuga",
               "Thuja/Juniperus", "Populus", "Acer", "Betula", "Fagus",
               "Ostrya/Carpinus", "Tilia", "Fraxinus", "Ulmus", "Quercus")

rownames(loss_table) <- paste0("*", sci_names, "*")

knitr::kable(loss_table)

```

**Supplemental Table 2**:  As Supplemental Table 1, except that in this analysis, all estimates are restricted to the subset of 8 km gridcells that had at least one observation in both the PLS and FIA datasets.  For most taxa, the percentages are similar (within a few percentage points) between Supplemental Tables 1 and 2, except for Quercus, which has a substantially lower estimate of percent area lost (ST1: 46% vs. ST2: 25.7%)  See also Supplemental Figure 1.

```{r, warning=FALSE, message=FALSE, echo=FALSE, results='as-is'}

if ('s2loss_table.rds' %in% list.files('data/output')) {
  s2loss_melt <- readRDS('data/output/s2loss_table.rds')
} else {

  s2loss_melt <- loss_plot(overlap = TRUE)
  saveRDS(s2loss_melt, file = 'data/output/s2loss_table.rds')
}

s2loss_table <- as.data.frame.matrix(table(s2loss_melt[[1]]$variable,
                                         s2loss_melt[[1]]$value))
s2loss_table <- data.frame(s2loss_table,
                         Modern = round((s2loss_table$Gain + s2loss_table$Presence) * .064, 0),
                         Historical = round((s2loss_table$Loss + s2loss_table$Presence ) * .064, 0))

s2loss_table[,1:3] <- round(analogue::tran(s2loss_table[,1:3], method = "proportion") * 100,1)

sci_names <- c("Larix", "Pinus", "Picea", "Abies", "Tsuga",
               "Thuja/Juniperus", "Populus", "Acer", "Betula", "Fagus",
               "Ostrya/Carpinus", "Tilia", "Fraxinus", "Ulmus", "Quercus")

rownames(s2loss_table) <- paste0("*", sci_names, "*")

knitr::kable(s2loss_table, justify = "left", style = "rmarkdown",
                     emphasize.rownames = FALSE)

```

**Supplemental Table 3**:  As Table 1 in the main text, except that the assessments of compounding (□), confounding (-), and counteracting (○) interactions are based on a historical climate normal that has been adjusted to account for 19th-century temperature trends reported from early meteorological time series at observing stations and US military forts (Burnette et al. 2010, Baker et al. 1985).  All entries in this table are identical to those in Table 1, indicating that the assessments of interaction strength are insensitive to this adjustment to mid-19th-century climates. See Supplemental Figure 2 for further details about this analysis.

```{r, warning=FALSE, message=FALSE, echo=FALSE, results = 'as-is'}

conf_pls <- readRDS('data/output/conf_pls.RDS')

conf_pls$taxa <- paste0("*", sci_names, "*")
colnames(conf_pls) <- c("Taxon", "$P_{ann}$", "$T_{max}$", "$T_{diff}$", "$T_{min}$")

knitr::kable(conf_pls)
```

**Supplemental Table 4**: As Table 1 in the main text, except that the compounding (□), confounding (-), and counteracting (○) assessments are based on a modern climate normal for 1950-1979 instead of 1990-2014. The results for ($T_{min}$), ($T_{diff}$), and ($P_{ann}$) are unchanged. For ($T_{max}$), the sign of the climate change flips from negative (cooling, $-_c$) to non-significant ($._c$), indicating that the apparent directionality of ($T_{max}$) change is sensitive to choice of modern climate normal.  (Note that directionality of ($T_{min}$) is unchanged, consistent with 20^th^- and 21^st^-century warming).   See also Supplemental Figures 3, 4.

```{r, warning=FALSE, message=FALSE, echo=FALSE, results = 'as-is'}

conf_fia <- readRDS('data/output/conf_fia.RDS')

conf_fia$taxa <- paste0("*", sci_names, "*")
colnames(conf_fia) <- c("Taxon", "$P_{ann}$", "$T_{max}$", "$T_{diff}$", "$T_{min}$")

knitr::kable(conf_fia)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

s2loss_melt[[2]]

```

# Supplemental Figures

**Supplemental Figure 1**:  As Figure 3 in the main text, except that all estimates of gains, losses, and presences are restricted to the subset of gridcells that had observations in both the PLS and FIA datasets.  See also Supplemental Table 2.

![](figures/boxplot_corrected_pls.png)

**Supplemental Figure 2**:  As Figure 4 in the main text, but with an adjustment to check the effects of the decision to use PRISM 1895-1924 climate normals instead of climate normals from the middle 19th century (i.e. a climate normal closer to the time of the PLS surveys).  No gridded instrumental temperature reconstructions are available in the upper Midwest prior to 1895, but early meteorological records from US military forts indicate 19th-century warming, with estimates of  warming at Fort Leavenworth, Kansas of 0.10oC decade^-1^ in the winter and 0.07oC decade^-1^ in the summer (Burnette et al. 2010).  Given this and the ~50 year offset between PLS observations and early PRISM data, the historical climate estimates used here may be biased by approximately +0.35^o^C for ($T_{max}$), +0.5^o^C for ($T_{min}$), and +0.15^o^C for t_diff. This would imply that the reported regional increase in ($T_{min}$) between historical to modern periods (3.3^o^C, Fig. 1) should be further increased to 3.8^o^C, the reported regional decrease in ($T_{max}$)  (-0.8^o^C) should be reduced to -0.45^o^C), and slightly larger reduction in $\Delta t_{diff}$ (-4.25^o^C). Hence, using a 1895-2024 climate normal may cause our estimates of historic climate change to be too conservative for ($T_{min}$) and ($T_{diff}$) and too large for ($T_{max}$).  The patterns shown in Supplementary Figure 3 are qualitatively unchanged from those shown in Fig. 4, suggesting that the findings of this paper are largely insensitive to this adjustment to account for 19th-century rises in temperature.

![](figures/boxplot_corrected_fia.png)

**Supplemental Figure 3**:  As Figure 4 in the main text, except that the modern climate normal is for 1950-1979 instead of 1990-2014. The results for ($T_{min}$) and ($P_{ann}$) are largely unchanged between this analysis and that shown in Fig. 4, indicating that the finding of tree-climate distribution shifts is robust to choice of modern cliamte normal.  However, ($T_{max}$) shows much closer overlap between the PLS and FIA periods, suggesting that the reported difference in tree-climate distributions for ($T_{max}$) is sensitive to choice of modern climate normal.

![](figures/hellinger_pls.png)

**Supplemental Figure 4**:  As Figure 5 in the main text, except that the modern climate normal is for 1950-1979 instead of 1990-2014.  Few differences are apparent between this figure and Figure 5, suggesting that the attribution of tree-climate distribution shifts to changes in climate versus changes in vegetation is insensitive to choice of modern climate normal.

![](figures/hellinger_fia.png)